---
title: "UW-Madison [Biomedical Data Science](https://bit.ly/MadBDS)"
author: "[Karl Broman](http://kbroman.org)"
date: "[`kbroman.org`](http://kbroman.org) <br/> [`github.com/kbroman`](https://github.com/kbroman) <br/> [`@kwbroman`](https://twitter.com/kwbroman) <br/> `broman@wisc.edu`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "kbroman_xaringan.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: '%current%'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message=FALSE, warning=FALSE,
                      echo=FALSE, include=FALSE,
                      fig.width=11.5, fig.height=6.5,
                      dev="svg", dev.args=list(pointsize=18))
```

# Why UW-Madison?

- Science
- People
- Madison

---

# Goal

Skills and training to

- Gain understanding of some area of biomedical research
- Develop statistical/computational methods <br/>
  to answer relevant scientific questions
- Implement those methods in software


---

# Why _biomedical data science_&nbsp;?

- growth in scale and complexity of biomedical data
- biostatistics + biomedical informatics

---

# About the BDS MS program

- 4 core courses
  - biostatistics
  - bioinformatics
  - medical image analysis
  - health informatics
- concentration electives (6 credits)
- data science electives (6 credits)
- research electives
  - research ethics (1 credit)
  - research project

---

# About the BDS PhD program

- flexible coursework
- but both statistics and computer science
- at least 2 biology courses
- 3 semester-long rotations
- literature course; professional skills course
- no written exam
- oral exam in 3rd year
- PhD thesis and oral defense

---

# PhD coursework: statistical genomics

```{r load_data}
courses <- read.csv("course_sequences.csv", stringsAsFactors=FALSE)
courses <- split(courses, courses$student)
```

```{r functions}
library(broman)
plot_semester <-
    function(data, x=0, y=0, width=100, heightper=100/13,
             color=broman::brocolors("crayons")[c("Magic Mint", "Blue Bell", "Melon",
             "Wild Watermelon", "Timberwolf", "Maize", "Wisteria")],
             add=FALSE)
{

    if(!add)
        plot(0,0,type="n", xlab="", ylab="", xaxt="n", yaxt="n",
             xlim=c(0,100), ylim=c(100, 0))

    text <- data$course
    color <- color[match(data$type, c("bio", "core", "elec", "eth", "prereq", "rot", "lit"))]

    ht <- heightper*data$credits
    y <- y + cumsum(c(0,ht[-length(ht)]))

    x <- rep(x, length(y))
    for(i in seq(along=y)) {
        rect(x, y[i], x+width, y[i]+ht[i], col=color[i], border="black")
        text(x+width/2, y[i]+ht[i]/2, text[i], cex=0.8)
    }
}

plot_student <-
    function(data, width=100, heightper=100/13, gapx = 10, gapy=heightper*2, legendw=20,
             color=broman::brocolors("crayons")[c("Magic Mint", "Blue Bell", "Melon",
             "Wild Watermelon", "Timberwolf", "Maize", "Wisteria")])
{
    # max per semester by year
    cr <- tapply(data$credits, data$semester, sum)
    cr <- matrix(cr, ncol=2, byrow=TRUE)
    cr <- apply(cr, 1, max)

    # start locations
    x <- c(0, width+gapx)
    y <- c(0, gapy+heightper*cr[1])
    y <- c(y, y[2]+gapy+heightper*cr[2])
    totalw <- max(x)+width+gapx*1.5+legendw
    totalh <- max(y)+heightper*cr[3]

    par(bty="n", mar=rep(0.1, 4))
    plot(0,0,type="n", xlab="", ylab="", xaxt="n", yaxt="n",
         xlim=c(-15,totalw+5), ylim=c(totalh+5, -15), xaxs="i", yaxs="i")
    px <- c(1,2,1,2,1,2)
    py <- c(1,1,2,2,3,3)
    for(i in 1:6) {
        plot_semester(data[data$semester==i,], x[px[i]], y[py[i]],
                      width, heightper, color, TRUE)
    }

    o <- c(5,2,6,3,1,7,4)
    legend(totalw-legendw-gapx, y[1], pch=22, pt.bg=color[o], col="black", pt.cex=3,
           c("Biology", "Core", "Elective", "Ethics", "Prereq", "Rotation", "Seminar")[o])

    text(x+width/2, rep(-7.5,2), c("Fall", "Spring"))
    text(rep(-7.5,3), y + cr*heightper/2, paste("Year", 1:3), srt=90)

}
```

```{r plot_genomics_student, include=TRUE}
plot_student(courses[[1]])
```

---

# PhD coursework: bioinformatics

```{r plot_bioinformatics_student, include=TRUE}
plot_student(courses[[2]])
```

---

# PhD coursework: clinical biostat/informatics

```{r plot_clin_biostat_student, include=TRUE}
plot_student(courses[[3]])
```
